---
title: "GLMM Analysis Report"
output: html_document
params:
  model: NULL
  anova: NULL
  formula: ""
  engine: ""
  family: ""
  dharma: NULL
  singular_fit: NULL
  convergence: NULL
  n_subjects: NULL
  n_observations: NULL
  fit_warnings: NULL
  session_info: NULL
---

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
library(knitr)
library(kableExtra)

# Render a data.frame as a styled HTML table and write it straight into the
# document (for use in results='asis' chunks). format = "html" is always set
# explicitly: with the default (markdown/pipe) format, kable escapes literal
# "|" characters - e.g. in "Pr(>|t|)" - as "&#124;", which then shows up
# verbatim in the report.
html_table <- function(df, caption = NULL, row.names = FALSE, align = NULL) {
  tbl <- knitr::kable(df, format = "html", caption = caption,
                      row.names = row.names, align = align)
  tbl <- kableExtra::kable_styling(
    tbl,
    bootstrap_options = c("striped", "hover", "condensed"),
    full_width = FALSE
  )
  cat(as.character(tbl), "\n")
  invisible(tbl)
}

# Evaluate an expression, returning NULL instead of an error/warning/message.
quiet <- function(expr) {
  tryCatch(suppressWarnings(suppressMessages(expr)), error = function(e) NULL)
}

# Tidy random-effects table with BOTH variance and standard deviation, for
# lme4 (merMod) and nlme (lme) models.
extract_random_effects <- function(model) {

  if (inherits(model, "merMod")) {
    vc <- as.data.frame(lme4::VarCorr(model))
    is_var <- is.na(vc$var2)

    components <- data.frame(
      Group      = as.character(vc$grp[is_var]),
      Effect     = ifelse(is.na(vc$var1[is_var]), "(Residual)", as.character(vc$var1[is_var])),
      Variance   = signif(vc$vcov[is_var], 4),
      `Std.Dev.` = signif(vc$sdcor[is_var], 4),
      check.names = FALSE, stringsAsFactors = FALSE
    )

    correlations <- NULL
    if (any(!is_var)) {
      correlations <- data.frame(
        Group       = as.character(vc$grp[!is_var]),
        Effects     = paste(vc$var1[!is_var], vc$var2[!is_var], sep = " ~ "),
        Covariance  = signif(vc$vcov[!is_var], 4),
        Correlation = signif(vc$sdcor[!is_var], 4),
        stringsAsFactors = FALSE
      )
    }
    return(list(components = components, correlations = correlations))
  }

  if (inherits(model, "lme")) {
    # nlme::VarCorr() returns a character matrix with "Variance"/"StdDev"
    # columns, plus group header rows such as "Subject = pdLogChol(1)".
    vc <- unclass(nlme::VarCorr(model))
    rn <- trimws(rownames(vc))
    has_corr <- "Corr" %in% colnames(vc)
    cur_group <- NA_character_
    rows <- list()

    for (i in seq_along(rn)) {
      if (grepl("=", rn[i], fixed = TRUE)) {
        cur_group <- trimws(sub("=.*$", "", rn[i]))
        next
      }
      variance <- suppressWarnings(as.numeric(vc[i, "Variance"]))
      if (is.na(variance)) next

      row <- data.frame(
        Group      = if (identical(rn[i], "Residual")) "Residual" else cur_group,
        Effect     = rn[i],
        Variance   = signif(variance, 4),
        `Std.Dev.` = signif(suppressWarnings(as.numeric(vc[i, "StdDev"])), 4),
        check.names = FALSE, stringsAsFactors = FALSE
      )
      if (has_corr) row$Corr <- trimws(vc[i, "Corr"])
      rows[[length(rows) + 1]] <- row
    }

    if (length(rows) == 0) return(NULL)
    components <- do.call(rbind, rows)
    if (has_corr && all(is.na(components$Corr) | components$Corr == "")) {
      components$Corr <- NULL
    }
    return(list(components = components, correlations = NULL))
  }

  NULL
}
```

## Methods

Model was fitted using **`r params$engine`**.

Family: **`r params$family`**

Formula: **`r params$formula`**

```{r sample-size, echo=FALSE, results='asis'}
if (!is.null(params$n_subjects) || !is.null(params$n_observations)) {
  cat(
    "Sample size used in this fit: **",
    ifelse(is.null(params$n_subjects) || is.na(params$n_subjects), "NA", params$n_subjects),
    " subjects**, **",
    ifelse(is.null(params$n_observations) || is.na(params$n_observations), "NA", params$n_observations),
    " observations**.\n",
    sep = ""
  )
}
```

---

## Model Fitting Diagnostics

This section reports diagnostics captured automatically at fit time and should be
reviewed before interpreting the results below. A singular fit, a non-convergence
message, or fitting warnings do not necessarily invalidate a model, but they do mean
its estimates (particularly variance components) should be interpreted with caution.

```{r diagnostics-table, echo=FALSE, results='asis'}
diag_rows <- list()

if (!is.null(params$singular_fit)) {
  diag_rows[["Singular fit"]] <- if (isTRUE(params$singular_fit)) {
    "YES \u26a0\ufe0f"
  } else if (identical(params$singular_fit, FALSE)) {
    "NO"
  } else {
    "N/A (not applicable to this engine, e.g. nlme)"
  }
}

if (!is.null(params$convergence)) {
  diag_rows[["Convergence"]] <- params$convergence
}

if (length(diag_rows) > 0) {
  diag_df <- data.frame(
    Check = names(diag_rows),
    Result = unlist(diag_rows, use.names = FALSE),
    stringsAsFactors = FALSE
  )
  html_table(diag_df, align = "l")
} else {
  cat("_No singularity/convergence information was supplied for this model._\n")
}
```

### Warnings raised during model fitting

```{r fit-warnings, echo=FALSE, results='asis'}
if (!is.null(params$fit_warnings) && length(params$fit_warnings) > 0) {
  cat("<ul>\n")
  for (w in params$fit_warnings) {
    cat("<li>", htmltools::htmlEscape(w), "</li>\n", sep = "")
  }
  cat("</ul>\n")
} else {
  cat("_No warnings were raised while fitting this model._\n")
}
```

---

## Results

### Model Summary

Full summary output of the fitted model, as printed by R.

```{r model-summary, echo=FALSE, message=FALSE, warning=FALSE}
old_opts <- options(width = 110)
tryCatch(
  print(summary(params$model)),
  error = function(e) cat("Model summary not available for this model:", conditionMessage(e), "\n")
)
options(old_opts)
```

### Random Effects

Both the **variance** (σ²) and the **standard deviation** (σ) of each random-effect
term are shown. The residual row is the within-group (observation-level) variance.

```{r random-effects, echo=FALSE, results='asis'}
re <- tryCatch(extract_random_effects(params$model), error = function(e) NULL)

if (!is.null(re) && nrow(re$components) > 0) {
  html_table(re$components, caption = "Random-effects variance components")

  if (!is.null(re$correlations)) {
    html_table(re$correlations, caption = "Random-effects covariances and correlations")
  }

  if (inherits(params$model, "glmerMod")) {
    cat("\n\n_For non-Gaussian GLMMs the random-effects variances are on the link",
        "(latent) scale, and there is no separate residual variance._\n")
  }

  # nlme only: residual correlation (correlation=) and variance (weights=) structures
  if (inherits(params$model, "lme")) {
    describe_struct <- function(s, label) {
      if (is.null(s)) return(invisible(NULL))
      pars <- tryCatch(coef(s, unconstrained = FALSE), error = function(e) NULL)
      pstr <- ""
      if (length(pars) > 0) {
        nm <- names(pars)
        if (is.null(nm)) nm <- seq_along(pars)
        pstr <- paste0(" (", paste(nm, signif(pars, 4), sep = " = ", collapse = ", "), ")")
      }
      cat("\n\n**", label, ":** `", class(s)[1], "`", pstr, "\n", sep = "")
    }
    describe_struct(params$model$modelStruct$corStruct, "Residual correlation structure")
    describe_struct(params$model$modelStruct$varStruct, "Residual variance structure (weights)")
  }
} else {
  cat("_Random-effects variance components could not be extracted for this model._\n")
}
```

### Fixed Effects

```{r fixed-effects, results='asis', echo=FALSE}
# nlme: summary(model)$tTable holds the fixed-effects table. Do NOT use
# coef(summary(model)) here - for an lme object that dispatches to
# coef.lme and returns per-subject coefficients instead.
# lme4: coef(summary(model)) returns the fixed-effects matrix.
coef_table <- if (inherits(params$model, "lme")) {
  tryCatch(summary(params$model)$tTable, error = function(e) NULL)
} else {
  tryCatch(coef(summary(params$model)), error = function(e) NULL)
}

if (is.null(coef_table)) {
  coef_table <- tryCatch(summary(params$model)$coefficients, error = function(e) NULL)
}

if (is.null(coef_table)) {
  cat("_Fixed-effects table not available for this model._\n")
} else {

  coef_df <- as.data.frame(coef_table, stringsAsFactors = FALSE)

  # Standardise nlme column names to the lme4 names.
  # nlme tTable columns: Value, Std.Error, DF, t-value, p-value
  nlme_names <- c(
    "Value"     = "Estimate",
    "Std.Error" = "Std. Error",
    "DF"        = "df",
    "t-value"   = "t value",
    "p-value"   = "Pr(>|t|)"
  )
  for (old_name in intersect(names(nlme_names), names(coef_df))) {
    names(coef_df)[names(coef_df) == old_name] <- nlme_names[[old_name]]
  }

  # Round numeric columns for readability
  if ("Estimate" %in% names(coef_df))    coef_df$Estimate     <- round(coef_df$Estimate, 3)
  if ("Std. Error" %in% names(coef_df))  coef_df$`Std. Error` <- round(coef_df$`Std. Error`, 3)
  if ("df" %in% names(coef_df))          coef_df$df           <- round(coef_df$df, 0)
  if ("t value" %in% names(coef_df))     coef_df$`t value`    <- round(coef_df$`t value`, 3)
  if ("z value" %in% names(coef_df))     coef_df$`z value`    <- round(coef_df$`z value`, 3)

  # Keep p-values numeric long enough to compute significance stars, then
  # format for display.
  p_col <- intersect(c("Pr(>|t|)", "Pr(>|z|)"), names(coef_df))

  if (length(p_col) == 1) {
    p_numeric <- suppressWarnings(as.numeric(coef_df[[p_col]]))

    sig_stars <- cut(
      p_numeric,
      breaks = c(-Inf, 0.001, 0.01, 0.05, 0.1, Inf),
      labels = c("***", "**", "*", ".", "")
    )

    coef_df[[p_col]] <- format.pval(p_numeric, digits = 3, eps = 0.0001)
    coef_df$" " <- sig_stars
  }

  html_table(coef_df, caption = "Fixed Effects", row.names = TRUE)
}
```

### ANOVA (Type III)

```{r anova, echo=FALSE}
if (!is.null(params$anova)) {
  print(params$anova)
} else {
  cat("ANOVA table not available for this model.\n")
}
```

### Model Performance

```{r performance, echo=FALSE, results='asis'}
mdl <- params$model

r2   <- quiet(performance::r2_nakagawa(mdl))
icc  <- quiet(performance::icc(mdl))
perf <- quiet(performance::model_performance(mdl))

# Safely pull a single numeric value out of a list / data.frame (or NULL)
pull <- function(x, nm) {
  if (!is.null(x) && !is.null(x[[nm]])) as.numeric(x[[nm]])[1] else NULL
}

add_metric <- function(df, label, value) {
  if (is.null(value) || length(value) == 0 || !is.finite(value[1])) return(df)
  rbind(df, data.frame(Metric = label, Value = signif(as.numeric(value[1]), 4),
                       stringsAsFactors = FALSE))
}

r2_marginal    <- pull(r2, "R2_marginal")
r2_conditional <- pull(r2, "R2_conditional")
if (is.null(r2_marginal))    r2_marginal    <- pull(perf, "R2_marginal")
if (is.null(r2_conditional)) r2_conditional <- pull(perf, "R2_conditional")

icc_adjusted   <- pull(icc, "ICC_adjusted")
icc_unadjusted <- pull(icc, "ICC_unadjusted")
if (is.null(icc_adjusted)) icc_adjusted <- pull(perf, "ICC")

var_df <- data.frame(Metric = character(), Value = numeric(), stringsAsFactors = FALSE)
var_df <- add_metric(var_df, "R2 marginal (fixed effects only)", r2_marginal)
var_df <- add_metric(var_df, "R2 conditional (fixed + random effects)", r2_conditional)
var_df <- add_metric(var_df, "ICC (adjusted)", icc_adjusted)
var_df <- add_metric(var_df, "ICC (unadjusted)", icc_unadjusted)

if (nrow(var_df) > 0) {
  html_table(var_df, caption = "Variance explained and clustering (R2, ICC)")
} else {
  cat("_R2 and ICC could not be computed for this model (for example, a model",
      "with no random effects, or a model class not supported by the performance",
      "package)._\n")
}

# Remaining fit indices (AIC, AICc, BIC, RMSE, Sigma, ...)
if (!is.null(perf)) {
  p <- as.data.frame(perf)
  p <- p[, setdiff(names(p), c("R2", "R2_conditional", "R2_marginal", "ICC")), drop = FALSE]
  if (ncol(p) > 0) {
    fit_df <- data.frame(
      Metric = names(p),
      Value  = unname(vapply(p, function(x) signif(suppressWarnings(as.numeric(x[1])), 4), numeric(1))),
      stringsAsFactors = FALSE
    )
    fit_df <- fit_df[is.finite(fit_df$Value), , drop = FALSE]
    if (nrow(fit_df) > 0) {
      html_table(fit_df, caption = "Information criteria and error measures")
    }
  }
}

is_reml <- tryCatch({
  if (inherits(mdl, "merMod")) lme4::isREML(mdl)
  else if (inherits(mdl, "lme")) identical(mdl$method, "REML")
  else FALSE
}, error = function(e) FALSE)

cat("\n\n**How to read these measures**\n\n")
cat("- **R2 marginal**: proportion of variance explained by the fixed effects alone.\n")
cat("- **R2 conditional**: proportion of variance explained by the fixed and random effects together.\n")
cat("- **ICC**: proportion of the total (non-fixed-effect) variance attributable to differences between groups/subjects; higher values mean observations within the same group are more alike.\n")
if (inherits(mdl, "glmerMod")) {
  cat("- For non-Gaussian GLMMs, R2 and ICC are computed on the latent (link) scale, so they are not directly comparable to Gaussian-model values.\n")
}
if (is_reml) {
  cat("- This model was fitted by REML: AIC/BIC are only comparable between models with identical fixed effects. Refit by ML to compare models with different fixed effects.\n")
}
```

### DHARMa Residual Diagnostics

```{r dharma, echo=FALSE, fig.width=8, fig.height=5}
if (!is.null(params$dharma)) {
  plot(params$dharma)
} else {
  cat("DHARMa residual diagnostics are only computed for lme4-based (glmerMod/lmerMod) models.\n")
}
```

---

## Reproducibility Information

<details>
<summary>R and package versions used to produce this report (click to expand)</summary>
```{r session-info, echo=FALSE}
if (!is.null(params$session_info)) {
  cat(paste(params$session_info, collapse = "\n"))
} else {
  cat("Session information was not supplied for this report.")
}
```
</details>
