---
title: "GLMM Analysis Report"
output: html_document
params:
  model: NULL
  anova: NULL
  formula: ""
  engine: ""
  family: ""
  dharma: NULL
  singular_fit: NULL
  convergence: NULL
  n_subjects: NULL
  n_observations: NULL
  fit_warnings: NULL
  session_info: NULL
---

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
library(knitr)
library(kableExtra)
```

## Methods

Model was fitted using **`r params$engine`**.

Family: **`r params$family`**

Formula: **`r params$formula`**

```{r sample-size, echo=FALSE, results='asis'}
if (!is.null(params$n_subjects) || !is.null(params$n_observations)) {
  cat(
    "Sample size used in this fit: **",
    ifelse(is.null(params$n_subjects) || is.na(params$n_subjects), "NA", params$n_subjects),
    " subjects**, **",
    ifelse(is.null(params$n_observations) || is.na(params$n_observations), "NA", params$n_observations),
    " observations**.\n",
    sep = ""
  )
}
```

---

## Model Fitting Diagnostics

This section reports diagnostics captured automatically at fit time and should be
reviewed before interpreting the results below. A singular fit, a non-convergence
message, or fitting warnings do not necessarily invalidate a model, but they do mean
its estimates (particularly variance components) should be interpreted with caution.

```{r diagnostics-table, echo=FALSE, results='asis'}
diag_rows <- list()

if (!is.null(params$singular_fit)) {
  diag_rows[["Singular fit"]] <- if (isTRUE(params$singular_fit)) {
    "YES \u26a0\ufe0f"
  } else if (identical(params$singular_fit, FALSE)) {
    "NO"
  } else {
    "N/A (not applicable to this engine, e.g. nlme)"
  }
}

if (!is.null(params$convergence)) {
  diag_rows[["Convergence"]] <- params$convergence
}

if (length(diag_rows) > 0) {
  diag_df <- data.frame(
    Check = names(diag_rows),
    Result = unlist(diag_rows, use.names = FALSE),
    stringsAsFactors = FALSE
  )
  print(
    kable(diag_df, row.names = FALSE, align = "l") %>%
      kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                    full_width = FALSE)
  )
} else {
  cat("_No singularity/convergence information was supplied for this model._\n")
}
```

### Warnings raised during model fitting

```{r fit-warnings, echo=FALSE, results='asis'}
if (!is.null(params$fit_warnings) && length(params$fit_warnings) > 0) {
  cat("<ul>\n")
  for (w in params$fit_warnings) {
    cat("<li>", htmltools::htmlEscape(w), "</li>\n", sep = "")
  }
  cat("</ul>\n")
} else {
  cat("_No warnings were raised while fitting this model._\n")
}
```

---

## Results

```{r model-summary, echo=FALSE, message=FALSE, warning=FALSE}
model_summary <- summary(params$model)
```

### Random Effects
```{r random-effects, echo=FALSE}
if (!is.null(model_summary$varcor)) {
  print(model_summary$varcor)
} else {
  cat("Random-effects variance-covariance summary not available for this model.\n")
}
```

### Fixed Effects
```{r fixed-effects, results='asis', echo=FALSE}
# Extract the fixed-effects coefficient table in a way that works for both
# lme4 and nlme models.  In particular, summary.lme() stores the fixed
# effects in a structure that differs from lme4's summary objects.
coef_table <- tryCatch(
  coef(summary(params$model)),
  error = function(e) NULL
)

if (is.null(coef_table)) {
  coef_table <- model_summary$coefficients
}

coef_df <- as.data.frame(coef_table, stringsAsFactors = FALSE)

# Standardise nlme column names to the names used by lme4 where possible.
# nlme::summary.lme() typically returns: Value, Std.Error, DF, t-value, p-value.
nlme_names <- c(
  "Value" = "Estimate",
  "Std.Error" = "Std. Error",
  "DF" = "df",
  "t-value" = "t value",
  "p-value" = "Pr"
)
for (old_name in intersect(names(nlme_names), names(coef_df))) {
  names(coef_df)[names(coef_df) == old_name] <- nlme_names[[old_name]]
}

# Round numeric columns for readability
if ("Estimate" %in% names(coef_df))    coef_df$Estimate    <- round(coef_df$Estimate, 3)
if ("Std. Error" %in% names(coef_df))  coef_df$`Std. Error` <- round(coef_df$`Std. Error`, 3)
if ("df" %in% names(coef_df))          coef_df$df          <- round(coef_df$df, 0)
if ("t value" %in% names(coef_df))     coef_df$`t value`   <- round(coef_df$`t value`, 3)

# Identify the p-value column. Keep the numeric values long enough to compute
# significance stars, then format them only for display.
 p_col <- intersect(c("Pr(>|t|)", "Pr(>|z|)"), names(coef_df))

if (length(p_col) == 1) {
  p_numeric <- suppressWarnings(as.numeric(coef_df[[p_col]]))

  sig_stars <- cut(
    p_numeric,
    breaks = c(-Inf, 0.001, 0.01, 0.05, 0.1, Inf),
    labels = c("***", "**", "*", ".", "")
  )

  coef_df[[p_col]] <- format.pval(
    p_numeric,
    digits = 3,
    eps = 0.0001
  )
  coef_df$" " <- sig_stars
}

kable(coef_df, caption = "Fixed Effects", row.names = TRUE, align = "r", escape = FALSE) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                full_width = FALSE)
```

### ANOVA (Type III)
```{r anova, echo=FALSE}
if (!is.null(params$anova)) {
  print(params$anova)
} else {
  cat("ANOVA table not available for this model.\n")
}
```

### DHARMa Residual Diagnostics
```{r dharma, echo=FALSE, fig.width=8, fig.height=5}
if (!is.null(params$dharma)) {
  plot(params$dharma)
} else {
  cat("DHARMa residual diagnostics are only computed for lme4-based (glmerMod/lmerMod) models.\n")
}
```

---

## Reproducibility Information

<details>
<summary>R and package versions used to produce this report (click to expand)</summary>
```{r session-info, echo=FALSE}
if (!is.null(params$session_info)) {
  cat(paste(params$session_info, collapse = "\n"))
} else {
  cat("Session information was not supplied for this report.")
}
```
</details>
